---
title: "Bourzikas_Unit 3 Homework"
author: "Grant Bourzikas"
date: "9/17/2018"
assignment:
output: html_document
---

Setup R Enviornment
```{r setup, include=TRUE}
knitr::opts_chunk$set(echo = TRUE)
library(plyr)
```

## Question 1.  Get Cloning
```{r include=TRUE}

# ls
# cd OneDrive\ -\ McAfee/Documents/SMU/6306\ -\ DDS\ -\ F18/Unit\ 3/
# mkdir APD
# cd APD
# git clone https://github.com/caesar0301/awesome-public-datasets""
```

## Quetsion 1.  Data Summary
### A. Read Data In
```{r include=TRUE}
df <- read.csv("~/OneDrive - McAfee/Documents/SMU/6306 - DDS - F18/Unit 3/awesome-public-datasets/Datasets/titanic.csv")
```

### B. Count Males and Femals and Plot Frequency
```{r include=TRUE}
df_Sex <- table(df$Sex)
head(df_Sex)
op <- par(mar=c(4,15,2,1), las= 2)
barplot(df_Sex,
          main = "Males vs Females on the Titanic", 
          xlab="Males", 
          ylab="Females", 
          col='Firebrick')
```

### C. Use Apply Functions
```{r include=TRUE}
df2.df <- data.frame(df$Age, df$Fare, df$Survived)
df_Sapply <- sapply(df2.df,na.rm=TRUE, mean)
```

## 3.	Function Building 
### A. Summary Level Data on Sleep Data
```{r include=TRUE}
sleep <- read.csv("~/OneDrive - McAfee/Documents/SMU/6306 - DDS - F18/Unit 3/APD/sleep_data_01.csv")
summary(sleep)
mean(sleep$RSES, na.rm = TRUE)
sd(sleep$RSES, na.rm = TRUE)

```

### B. Dataframe Object Summary
```{r include=TRUE}
sleep_Med <-- median(sleep$Age, na.rm = TRUE)
sleep_Min <-- min(sleep$RSES, na.rm = TRUE)
sleep_Max <-- max(sleep$RSES, na.rm = TRUE)
sleep_mean <-- mean(sleep$Age, na.rm = TRUE)
sleep_sd <-- sd(sleep$RSES, na.rm = TRUE)
sleep_Range <-- range(sleep$Duration, na.rm = TRUE)
report <- data.frame(sleep_Med, sleep_mean/5, sleep_sd/5, sleep_Range)
round(report, digits=2)

```

## 4.	FiveThirtyEight Data
### A. Install fivethirtyeight package
```{r include=TRUE}
library(fivethirtyeight)
```

### B. Assign 22nd Dataset
```{r include=TRUE}
df <- college_recent_grads
```

### C. Comment to URL
```{r include=TRUE}
# https://fivethirtyeight.com/features/the-economic-guide-to-picking-a-college-major/
```

### D. Dimensions and column names
```{r include=TRUE}
dim(df)
names(df)
```


## Data Summary
### A. Column Names and Column Counts
```{r include=TRUE}
names(df)
ncol(df)
```

### B. Major Catagory
```{r include=TRUE}
major_count <- count(df, 'major_category')
major_count
```

### C. Bar Plot of Major Catagories
```{r include=TRUE}
op <- par(mar=c(4,15,2,1), las= 2)
barplot(
    height=major_count$freq,
    names.arg = major_count$major_category,
    horiz = TRUE,
    main = 'The  Guide To Picking A College Major',
    xlab = 'Frequency',
    ylab = 'Catagory',
    col = 'FireBrick')

```

### D. Write CSV
```{r include=TRUE}
write.csv(df, file = 'collegeGradStudentsF.csv',row.names = FALSE)
```

## Question 3 Codebook
https://github.com/gbourzikas/DDS6306/blob/master/README.md
